Kindly Bent to Free Us

### 2020-06-20 kindly/soa.tex

\begin{figure}[tp]
  \begin{center}
    \newcommand\YC{\color{Green}}
    \newcommand\NC{\color{Red}}
    \newcommand\MC{\color{Orange}}
    \newcommand\Y{{\YC{\ding{52}}}\xspace}
    \newcommand\N{{\NC{\ding{56}}}\xspace}
    \newcommand\M{\MC{\textasciitilde}} % "Meh"

@@

    \begin{tabular}{l|ccccccccc}
      \textbf{Language}
      & UAL & \rot{Borrows} & \rot{State}
      & \rot{Resource Subsumption}
      & \rot{Resource Polymorphism}
      & \rot{Identity}
      & \rot{Inference} & \rot{Formalisation}
      & \rot{Basis}
      \\\hline
      System F\degree~\citep{DBLP:conf/tldi/MazurakZZ10}
      & UL & --- & \N & \N & \N & \N & \N & \YC Coq & System F
      \\
      Alms~\citep{DBLP:conf/popl/TovP11}
      & UA & --- & \Y & \M & \Y & \Y & \MC Local & \N & ML \\
      Quill~\citep{DBLP:conf/icfp/Morris16}
      & UL & --- & \N & \N & \Y & \N & \YC Principal & \MC Manual & Qualified types\\
      Lin. Haskell~\citep{DBLP:journals/pacmpl/BernardyBNJS18}
      & UL & --- & \M & \N & \Y & \N & \MC Non-principal & \MC Manual & Haskell \\
      Mezzo~\citep{DBLP:phd/hal/Protzenko14}
      & UA & --- & \Y & \M & \Y & \Y & \MC Local & \YC Coq & --- \\
      \hline
      Rust~\citep{rust,DBLP:journals/pacmpl/0002JKD18}
      & UA & \Y & \Y & \Y & \M & \N & \MC Local & \YC Coq & --- \\
      Plaid~\citep{DBLP:conf/oopsla/AldrichSSS09}
      & UA & --- & \Y & \Y & \Y & \Y & \N & \MC Manual & Java \\
      \hline
      \lang
      & UAL & \Y & \Y & \Y & \Y & \N & \YC Principal & \MC Manual & ML/HM(X)
      \\
    \end{tabular}
  \end{center}
  \caption{Comparison matrix}
  \label{fig:comparison-matrix}
\end{figure}

@@

The comparison matrix in \cref{fig:comparison-matrix} gives an
overview over the systems discussed in this section. The column UAL
specifies the substructural features supported (unrestricted, affine,
linear). Borrows and state indicate support for the respective
feature. ``Resource Subsumption'' indicates that unrestricted elements 
can be promoted to affine and then linear. This promotion applies to
all objects, resources,
borrows, or closures. ``Resource Polymorphism'' refers to polymorphism over
substructural features, notably for closures.
``Identity'' indicates that the language supports a notion
of identity, usually through existential types, as described in
\cref{identity}.
Inference is self-explanatory, formalization
refers to the existence of a formal semantics and type soundness proof. Basis is self-explanatory.

>>>>

\newcommand\YC{\color{Green}}
\newcommand\NC{\color{Red}}
\newcommand\MC{\color{Orange}}
\newcommand\Y{{\YC{\ding{52}}}\xspace}
\newcommand\N{{\NC{\ding{56}}}\xspace}
\newcommand\M{{\MC{\textasciitilde}}\xspace} % "Meh"
\begin{figure}[tp]
  \begin{center}

@@

    \begin{tabular}{l|ccccccccc}
      \textbf{Language}
      & UAL & \rot{Borrows} & \rot{State}
      & \rot{Resource Subsumption}
      & \rot{Resource Polymorphism}
      & \rot{Identity}
      & \rot{Inference} & \rot{Formalisation}
      & \rot{Basis}
      \\\hline
      System F\degree~\citep{DBLP:conf/tldi/MazurakZZ10}
      & UL & \N & \N & \N & \N & \N & \N & \YC Coq & System F
      \\
      Alms~\citep{DBLP:conf/popl/TovP11}
      & UA & \N & \Y & \M & \Y & \Y & \MC Local & \N & ML \\
      Quill~\citep{DBLP:conf/icfp/Morris16}
      & UL & \N & \N & \N & \Y & \N & \YC Principal & \MC Manual & Qualified types\\
      Lin. Haskell~\citep{DBLP:journals/pacmpl/BernardyBNJS18}
      & UL & \N & \M & \N & \Y & \N & \MC Non-principal & \MC Manual & Haskell \\
      Mezzo~\citep{DBLP:phd/hal/Protzenko14}
      & UA & \M & \Y & \M & \Y & \Y & \MC Local & \YC Coq & --- \\
      \hline
      Rust~\citep{rust,DBLP:journals/pacmpl/0002JKD18}
      & UA & \Y & \Y & \Y & \M & \N & \MC Local & \YC Coq & --- \\
      Plaid~\citep{DBLP:conf/oopsla/AldrichSSS09}
      & UA & \N & \Y & \Y & \Y & \Y & \N & \MC Manual & Java \\
      \hline
      \lang
      & UAL & \Y & \Y & \Y & \Y & \N & \YC Principal & \MC Manual & ML/HM(X)
      \\
    \end{tabular}
  \end{center}
  \caption{Comparison matrix}
  \label{fig:comparison-matrix}
\end{figure}

@@

The comparison matrix in \cref{fig:comparison-matrix} gives an
overview over the systems discussed in this section.
Each column indicates if a feature is present (\Y), absent (\N) or
partially supported (\M) such as when the feature is limited or can
only be obtained through a non-trivial encoding.
The column UAL
specifies the substructural features supported (unrestricted, affine,
linear). Borrows and state indicate support for the respective
feature. ``Resource Subsumption'' indicates that unrestricted elements 
can be promoted to affine and then linear. This promotion applies to
all objects, resources,
borrows, or closures. ``Resource Polymorphism'' refers to polymorphism over
substructural features, notably for closures.
``Identity'' indicates that the language supports a notion
of identity, usually through existential types, as described in
\cref{identity}.
Inference is self-explanatory, formalization
refers to the existence of a formal semantics and type soundness proof. Basis is self-explanatory.
